Title

### 2012-12-04 inc/photocathode/plasmon/plasmon.tex
move photocathode topics into subdirectory

For proof-of-concept purposes the first experiment was run on a commercially available gold foil on glass holographic grating.
Like the Kretschmann geometry, a foil on glass grating is vulnerable to heating problems; a plasmon creates a large amount of heat but the glass is a poor heat conductor.
The grating was operational as the angles were tuned attempting to drive the plasmon.
At the instant that the plasmon started, a bright flash was visible on the scintillator screen then disappeared.

After this, the grating showed signs of damage.
\ref{fig:grating-damage} shows the results of Scanning Electron Microscopy (SEM) on the grating.
The damage pattern is round and of the size of the laser spot used to drive the plasmon.
The extent of the damage indicates a large temperature increase, which given the other evidence of the test, can be assumed to be from the sudden appearance of a surface plasmon.

The most important result was that a plasmon was driven successfully; this created the heat that destroyed the grating.
Further this experiment validated the decision not to persue a Kretschmann geometry for fear of heat damage at these high powers.

@@

>>>>

For proof-of-concept purposes the first experiment was run on a commercially available gold foil on glass holographic grating (750 lines/mm).
Like the Kretschmann geometry, a foil on glass grating is vulnerable to heating problems; a plasmon creates a large amount of heat but the glass is a poor heat conductor.
The aforementioned laser (emitting $\sim$200fs pulses of green light at 30MW/cm$^2$) was used to drive the plasmon.

The input laser angle was slowly changed until the plasmon oscillation is visible.
At the instant that the plasmon started, a bright flash was visible on the scintillator screen then disappeared.

After this, the grating showed visual signs of damage, both on the surface and in the reflected light pattern.
\ref{fig:grating-damage} shows the results of Scanning Electron Microscopy (SEM) on the grating.
The damage pattern is round and of the size of the laser spot used to drive the plasmon.
The extent of the damage indicates a large temperature increase, which given the other evidence of the test, can be assumed to be from the sudden appearance of a surface plasmon.

Though this conclusively shows that a plasmon was driven successfully, the heat generated by the plasmon oscillation destroyed the grating.
Further this experiment validated the decision not to pursue a Kretschmann geometry for fear of heat damage at these high powers.

@@

After the destruction of the gold foil on glass grating, new gratings were designed to have a higher thermal conductivity.
The second generation grating was constructed from a silicon substract then coated with a gold film.
The thermal conductivity of silicon is approximately 100 times that of glass. %TODO reference needed?

Collaborators at Argonne National Laboratory's Electron Microscopy Center, Material Science Division used a defocused Zeiss 1540XB focused ion beam (FIB) to iteratively mill an approximately sinusoidal shape into the surface of a silicon wafer.
Several regions were created each 100x100$\mu$m in area, having a $\sim$1$\mu$m period and $\sim$10\% modulation depth.
An SEM image of one such region is shown in \ref{fig:fib-si-sem}.

Using the Varian e-beam deposition chamber at the UIC Nano Core Facility (NCF), I deposited a thin layer of chromium as a binding agent, then a 300nm film of gold.
This thickness of gold is several times thicker than the skin depth of green light on gold, therefore the laser radiation cannot interact with the silicon.

\begin{figure}
  \centering
  \includegraphics{HighMagSEM.jpg}
  \caption{
    SEM image of sinusoidal grating on silicon.
    Pattern milled by Focused Ion Beam (FIB) milling at Argonne National Lab (ANL).
    Sample was later coated with 300 nm gold film.
  }
  \label{fig:fib-si-sem}
\end{figure}
